---
title: "Week 9, Day 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(PPBDS.data)
library(rstanarm)
library(tidyverse)

# We will use the `kenya` dataset from PPBDS.data. This dataset is from Harris,
# Kamindo and Van der Windt (2020): "Electoral Administration in Fledgling
# Democracies: Experimental Evidence from Kenya." The authors worked with
# Kenya's electoral commission in 1,674 communities by assigning polling
# stations to either a control group or to one of five other treatments.

# For simplicity we will only consider polling stations assigned to either
# "control" or to "local". The latter means a treatment in which election
# officials visited the community's polling station for two days, thereby making
# it easier for people to register.

# The outcome variable, `reg_chg`, is the change in voter registration in each
# community. `poverty` is the percentage of the community below the poverty
# level. `distance` is the number of kilometers between the polling station and
# the central political office (where registrations normally take place).
# `pop_density` is a measure of population density around the polling station.


week_9 <- kenya %>% 
  rename(reg_chg = reg_byrv13) %>% 
  filter(treatment %in% c("control", "local")) %>% 
  droplevels() %>% 
  select(reg_chg, treatment, poverty, 
         distance, pop_density)
```


## Scene 1

**Prompt:** In addition to the outcome variable and the treatment, we have 3 statistics for the communities in the sample --- poverty, distance and population density. See code comments above for details. Never hurts to do a little exploratory data analysis. Look at the data!

* Make a scatter plot of `poverty` on the y-axis and `distance` in the x-axis. Include a `geom_smooth()` with `method` set to "lm" and `se` equal to FALSE. Write a sentence to answer these questions:
```{r sc1}
week_9 %>%
  ggplot(aes(x = distance, y = poverty)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method = lm, se = FALSE) + 
  labs(title = "Poverty and Distance at Kenya Polling Stations",
       subtitle = 'Poorer communities are farther away from polling stations',
       x = 'Distance (km)',
       y = 'Poverty Measure',
       caption = 'Data from Harris, Kaminda, and Van der Wind (2020)')
  

```

  + Does poverty cause distance? Or does distance cause poverty? Or both? Or neither?
  
  Neither, because although there seems to be a positive correlation between
  poverty and distance, there's no causation.
  
  + How accurate is the model beyond 90 kilometers?  
  
  Not that accurate, because there's only one data point >90 km.
  
  + Eye-balling the chart, how different is the poverty rate between two polling stations whose distance differs by 50 kilometers?
  
  There seems to be a ~10% difference in poverty rate in polling stations 
  whose distance differs by 50 km.
  
* Create a linear model which relates poverty to distance. This is similar to the models we created last week which related age to party. See the discussion in Chapter 9 for advice.  Write down the math. 
  
  $$ poverty_i = \beta_0 + \beta_1 distance_i + \epsilon_i $$
  
* Fit a `stan_glm()` model which relates poverty to distance. Print a summary. (Hint: In order to see the coefficient of `distance`, you will probably need to use the `digits` argument to `print()`.) Interpret the meaning of the two important parameters.  

```{r sc1}
fit_obj <- stan_glm(data = week_9,
         poverty ~ distance,
         family = gaussian(),
         refresh = 0)

print(fit_obj, digits = 3)
```

The intercept indicates the poverty proportion when the distance from the polling
station is 0. The distance parameter tells us the coefficient of distance (i.e.
for every increase in 1 km, there's a 0.002 increase in the poverty proportion).


## Scene 2

**Prompt:** Let's make a model with `reg_chg` as the outcome variable and with treatment, poverty, distance and pop_density as the predictors. (Using outcome/predictors is the typical nomenclature in statistics. In Economics, we would refer to dependent/independent variables.)

* Provide the math for this model.

  $$ \textit{reg_chg}_i = \beta_1 x_{t,i} + \beta_2 x_{p,i} + \beta_3 x_{d,i} 
  + \beta_4 x_{pd,i} + \epsilon $$

* Estimate the model. Assign the fitted model to `fit_2`. Interpret the important parameters of the model.

```{r sc2}
fit_obj_2 <- stan_glm(data = week_9,
         reg_chg ~ treatment + poverty + distance + pop_density,
         refresh = 0)

print(fit_obj_2, digits = 3)
```

Poverty and 

Calculate confidence intervals for the difference between two communities, one with a poverty rate of 80% and one of 60%. Use posterior_predict to help you.



